Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
luis comments added + fixzinho na discussao

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.5\textwidth]{figures/language_dist.pdf}
 	\caption{Distribution of Security Refactorings per Language}
	\label{fig:lang}
\end{figure}

@@

The impact of each type of security refactoring on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. The results of the
Wilcoxon signed-rank test by pattern returned statistical evidence which
supports that the maintainability of open-source software decreases for 4
different patterns: \emph{Broken Autentication} (p = 0.018), \emph{Denial-of-Service}
(p = 0.029), \emph{Memory Leak} (p = 0.009) and \emph{Miscellaneous} (p = $3.07$x$10^{-4}$).
The results also returned statistical significance that evidence that maintainability
is more likely to not suffer differences for open-source software after removing
\emph{Components with Known Vulnerabilities} (p = $7.96$x$10^{-4}$) and fixing \emph{Cross-Site
Scripting Vulnerabilities} (p = 0.001). \emph{Injection} and \emph{Cross-Site Request Forgery}
produced more cases where the refactorings had a negative impact in the maintainability.
However, the results obtained for these patterns are not statistically significant.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Known Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
type of vulnerabilities are one line fixes --- adding an \texttt{escape}
function or changing the version of the component, respectively. Whereas,
\emph{Broken Authentication} and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.

>>>>

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.5\textwidth]{figures/language_dist.pdf}
 	\caption{Distribution of Security Refactorings per Language\Luis{Consegues ordenar estas linguagens da mesma forma que estão ordenadas na figura7}}
	\label{fig:lang}
\end{figure}

@@

The impact of each type of security refactoring on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. The results of the
Wilcoxon signed-rank test by pattern returned statistical evidence which
supports that the maintainability of open-source software decreases for 4
different patterns: \emph{Broken Autentication} (p = 0.018), \emph{Denial-of-Service}
(p = 0.029), \emph{Memory Leak} (p = 0.009) and \emph{Miscellaneous} (p = $3.07$x$10^{-4}$).
The results also returned statistical significance that evidence that maintainability
is more likely to not suffer differences for open-source software after removing
\emph{Components with Known Vulnerabilities}\Luis{o resultado e bem diferente dos regular commits. este refactoring tem alguma particularidade para deixar a maintainability intacta? convém falar sobre isso na discussão} (p = $7.96$x$10^{-4}$) and fixing \emph{Cross-Site
Scripting Vulnerabilities} (p = 0.001). \emph{Injection} and \emph{Cross-Site Request Forgery}
produced more cases where the refactorings had a negative impact in the maintainability.
However, the results obtained for these patterns are not statistically significant.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Known Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
types of vulnerabilities do not need extra lines to be fixed, as showed in
Listing~\ref{lst:fix}. 
Whereas, \emph{Broken Authentication} and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.
